---
title: "PA1_template.Rmd"
author: "Emily"
date: "7/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an RMarkdown document for Reproducible Research Course Project 1
First step is to load packages

```{r}
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(VIM)
library(mice)
set.seed(1010)
```

##Question 1: Load data/clean dates
```{r}
activity <- read_csv("activity.csv")
str(activity)
activity$date <- as.Date(activity$date)
```
##Question 2: Total steps taken per day
###Total Steps taken per day
```{r}
stepstotal <- aggregate(steps~date,activity,sum)
````
###Histogram of average steps per day
```{r}

hist(stepstotal$steps,xlab="Total steps",main="Total number of daily steps")

```
### Average Steps without missing
```{r}
stepsaverage <- mean(stepstotal$steps, na.rm=TRUE)
stepsaverage
````
### Median Steps without missing
```{r}
stepsmedian <- median(stepstotal$steps, na.rm=TRUE)
stepsmedian
```

##Question3: Average Daily Activity Pattern
###aggregating steps by interval and getting mean
```{r}
stepsinterval <- aggregate(steps~interval,activity,mean)
```

###Prep dataset for merging
```{r}
names(stepsinterval)[names(stepsinterval) == 'steps'] <- 'AverageSteps'
```

###time series plot of 5-minute interval and average steps taken
```{r}
plot(stepsinterval,type="l",xlab="Interval [5-min increments]",ylab="Average number of steps",main="Average number of steps per interval")
```

###what 5 minute interval contains the max steps?
```{r}
maxsteps <- stepsinterval$interval[which.max(stepsinterval$AverageSteps)]
maxsteps
```
##Question 4: Missing Values
###Calculate and report missing values
```{r}
missing <- sum(is.na(activity))
```

###Strategy for missing values... Going to impute
```{r}
imputing = merge(activity, stepsinterval, by="interval")
```
###Creating new dataset merged with imputed data
```{r}
imputing$steps[is.na(imputing$steps)] = imputing$AverageSteps[is.na(imputing$steps)]
```

###make a histogram
```{r}
newactivity <- aggregate(steps~interval, imputing,sum)

hist(newactivity$steps,xlab="Total steps",main="Total number of daily steps (Imputed)")
```
###Calculating mean of imputed data
```{r}
imputedmean <- mean(newactivity$steps, na.rm=TRUE)
imputedmean
```

###Calculating median of imputed data
```{r}
imputedmedian <- median(newactivity$steps, na.rm=TRUE)
imputedmedian
```
###difference between real and imputed means/medians
```{r}
meandiff <- imputedmean - stepsaverage
meandiff
```


```{r}
mediandiff <- imputedmedian - stepsmedian
mediandiff
```
###impact of missing values: Using imputed data brings down the average steps when compared with
###excluding the missing data


##Question 5: Difference on weekends?
###Create factor variable:weekend/weekday
```{r}
daytype <- function(date) {
    if (weekdays(date) %in% c("Saturday", "Sunday")) {
        "weekend"
    } else {
        "weekday"
    }
}
```
###apply to data
```{r}
imputing$daytype <- as.factor(sapply(imputing$date, daytype))
```
###make panel plot
```{r}
library(plyr)
imputedsteptype <- ddply(imputing,.(interval,daytype),summarize,stepsavg=mean(steps))
library(lattice)
xyplot(stepsavg~interval|daytype,data=imputedsteptype,type="l",layout = c(1,2),
       main="Average steps (imputed) per interval and type of day",
       ylab="Average number of steps",xlab="Interval [5-min increments]")
```












